The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-10 rational.tex
intro part 2

In the algebra of rational numbers {{1 = (1/a) * a}} if {{a}} is
nonzero.  In our framework this algebraic identity corresponds to a
type isomorphism {{1 <-> (1/a) * a}} (written alternately as
{{a-o*a}}) whose witness maybe thought of as the `identity function'
at the type {{a}}.  For example, for the datatype {{color = R|Gr|B}},
there are three possible witnesses for this isomorphism mapping {{()}}
to either {{(1/R,R)}} or {{(1/Gr,Gr)}} or {{(1/B,B)}} with the
particular witness determined by further constraints encountered
during the computation. More generally, the witness for the
isomorphism above may be represented as {{(1/alpha,alpha)}}
where~{{alpha}} is a fresh logic variable of type {{a}}. With this
view of {{a -o* a}} functions, let us consider the following
manipulations:

@@

\begin{itemize}
\item Using the fact that {{1}} is the multiplicative unit, generate
  from the input {{()}} the value {{((),())}} of type {{1 * 1}};
\item Apply the isomorphism {{1 <-> (1/a) * a}} in parallel to each of
  the components of the above tuple. The resulting value is
  {{((1/alpha1,alpha1),(1/alpha2,alpha2))}} where {{alpha1}}, and
  {{alpha2}} are fresh logic variables.
\item Using the fact that {{*}} is associative and commutative, we can
  rearrange the above tuple to produce the value:

@@

At this point we have constructed a strange mixing of two `{{a-o*a}}'
functions; `inputs' of one form `outputs' of the other. If
{{(1/alpha1,alpha2)}} is held by one subcomputation and
{{(1/alpha2,alpha1)}} is held by another subcomputation, these
`remixed functions' form a communication channel between the two
concurrent subcomputations -- unifying {{1/alpha1}} with {{color}}
`{{R}}' in one subcomputation, fixes {{alpha1}} in the other. The type
{{a}} thus takes the role of the type of the communication channel,
indicating how much information can be communicated between the two
subcomputations.  Depending on the choice of the type {{a}}, an
arbitrary number of bits may be communicated.

>>>>

In the algebra of rational numbers {{1 = (1/a) * a}} if {{a}} is
nonzero.  In our framework this algebraic identity corresponds to a
type isomorphism {{1 <-> (1/a) * a}} (written alternately as
{{a-o*a}}) whose witness maybe thought of as the `identity function'
at the type {{a}}.  For example, for the datatype {{color = R|Gr|B}},
there are three possible witnesses for this isomorphism mapping {{()}}
to either {{(1/R,R)}} or {{(1/Gr,Gr)}} or {{(1/B,B)}} with the
particular witness determined by further constraints encountered
during the computation. More generally, the witness for the
isomorphism above may be represented as {{(1/alpha,alpha)}}
where~{{alpha}} is a fresh logic variable of type {{a}}. With this
view of `{{a -o* a}}' functions, let us consider the following
manipulations:

@@

\begin{itemize}
\item Using the fact that {{1}} is the multiplicative unit, generate
  from the input {{()}} the value {{((),())}} of type {{1 * 1}};
\item Apply the isomorphism {{1 <-> (1/a) * a}} in parallel to each of
  the components of the above tuple. The resulting value is
  {{((1/alpha1,alpha1),(1/alpha2,alpha2))}} where {{alpha1}} and
  {{alpha2}} are fresh logic variables;
\item Using the fact that {{*}} is associative and commutative, we can
  rearrange the above tuple to produce the value:

@@

At this point we have constructed a strange mix of two `{{a-o*a}}'
functions; `inputs' of one function manifests as `outputs' of the
other. If {{(1/alpha1,alpha2)}} is held by one subcomputation and
{{(1/alpha2,alpha1)}} is held by another subcomputation, these
`remixed functions' form a communication channel between the two
concurrent subcomputations. Unifying {{1/alpha1}} with {{color}}
`{{R}}' in one subcomputation, fixes {{alpha1}} to be {{R}} in the
other. The type {{a}} thus takes the role of the type of the
communication channel, indicating how much information can be
communicated between the two subcomputations.  Depending on the choice
of the type {{a}}, an arbitrary number of bits may be communicated.
